A Switching Planner for Combined Task and Observation Planning

### 2011-02-08 subarchitectures/planner.sa/trunk/src/latex/AAAI11/evaluation.tex
changes requested by moritz

We evaluate our approaches in robot exploration tasks from home and
office environments. Spatially, these consist of {\em rooms}
(office/kitchen/etc), and an underlying topological map that models
smaller areas of space, called {\em places}, and connectivity between
those. The mobile {\em robot} and {\em visual objects} inhabit the
topological space, the latter indicate the category of space they
inhabit -- e.g., spoons are likely to be in kitchens. By examining
view cones at places for particular objects, the robot is able to: (1)
categorise space at high (room) and low (place) levels, and (2) find
objects for the user, exploiting information about object
co-occurrence and room categories for efficiency. Also, in the
presence of a person, the robot can ask them about the location of
objects.

We compare our switching planner with the {\em baseline} in several
realistic tasks, with the number of rooms ranging from 3 (12-places,
16-objects, $|$states$|$$>\sim10^{21}$) to 6 (26-places, 21-objects,
$|$states$|$ $>10^{36}$). For or experiments we model three sensors:
{\em Reliable} sensors have a $0.1$ probability of a false negative,
{\em semi-reliable} have a chance of $0.3$ of false negative and $0.1$
of false positive , and finally {\em noisy} sensors with probabilities
of $0.5$ and $0.2$ respectively. Each object class is assigned one
sensor model, so e.g. cornflakes may be harder to detect than
refrigeratos. To determine the influence of the sensor model, we
performed several experiments where we only changed the difficulty of
the target object(s). All non-target objects remain unchanged so the
planner may exploit object co-occurances with reliable objects to
compensate for the increased sensor noise.

We evaluate with DT sessions taking $b_0$ admitting between 20 and 200
abstract states with non-zero probability. We run 50 simulations in
each configuration, and have a timeout on each simulation of 30
minutes (1800 seconds)\footnote{All experiments were conducted on a
 2.66GHz Intel Xeon X5355 CPU using one CPU core.}. The continual planning
times are reported in Figure~\ref{fig:results-time}, and the quality
data in Figure~\ref{fig:results-quality}.
%%
For Items~$a$-$e$ the goal is to find objects and report their
position to a user. In some cases there is a non-zero probability that
no plan exists, as the desired object might not be present in the
environment. Item~$f$ reports for tasks requiring indirect sensing,
where the robot must relocate to a room with a particular category
(e.g. kitchen). the {\em baseline} is not able to complete Item~$f$,
so is omitted from that reporting.  Because reward is only allocated
once, on the achievement of the goal,  we find it
intuitive to report average plan costs and the success rates in
problems that admit a solution (i.e., positive reward scaled by a
constant factor).

@@

\begin{figure}[h!]
  % \centering
  % \includegraphics{dora1-quality}\hfill
  % \vspace{2mm}
  \includegraphics{pomdp-quality}\hfill
  \includegraphics{pomdp-solvable-quality}\hfill
  \includegraphics{dora2-quality}\hfill
  % \vspace{2mm}
  \includegraphics{dora3-quality}\hfill
  % \vspace{2mm}
  \includegraphics{dora4-quality}\hfill
  \vspace{2mm}
  \includegraphics{dora56-quality}\hfill
  % \vspace{2mm}
  % \includegraphics{dora-cat-quality}\hfill
  \caption{Average plan costs and number of successful runs.}
  \label{fig:results-quality}
\end{figure}

We find that if sensing is reliable there is little to be gained
through DT sessions, as the greedy approach of the {\em baseline} is
sufficient. As sensing degrades contingent planning pays off.  Time
spent in DT planning increases steeply as the abstraction $\bstate_0$
becomes more refined.  That refinement seems to be paying off in terms
of the success rate, particularly for tasks $d$ and $e$. For less
refined initial configurations, the increased cost of DT
planning is compensated for by a decrease in Fast Downward planning
times. The relatively high success rate irrespective of the level of
refinement in the initial configuration indicates the effectiveness of
using conditional entropy to guide abstraction refinement in our
setting.

>>>>

We evaluate our approaches in robot exploration tasks from home and
office environments. Spatially, these consist of {\em rooms}
(office/kitchen/etc), and an underlying topological map that models
smaller areas of space, called {\em places}, and connectivity between
those. The mobile {\em robot} and {\em visual objects} inhabit the
topological space, the latter indicate the category of space they
inhabit -- e.g., spoons are likely to be in kitchens. By examining
view cones at places for particular objects, the robot is able to: (1)
categorise space at high (room) and low (place) levels, and (2) find
objects for the user, exploiting information about object
co-occurrence and room categories for efficiency. Also, in the
presence of a person, the robot can ask them about the category of the
current room.

We compare switching with the {\em baseline} in several realistic
tasks, with the number of rooms ranging from 3 (12-places, 16-objects,
$|$states$|$ $>10^{21}$) to 6 (26-places, 21-objects, $|$states$|$
$>10^{36}$). We also compare those systems with an approximately
optimal policy computed using Smith's {\sc zmdp} for small 2 room
problems (4-places, 3-objects, $|$states$|$ $\simeq 5000$). For or
experiments we model three sensors: {\em Reliable} sensors have a
$0.1$ probability of a false negative, {\em semi-reliable} have a
chance of $0.3$ of false negative and $0.1$ of false positive , and
finally {\em noisy} sensors with probabilities of $0.5$ and $0.2$
respectively. Each object class is assigned one sensor model, so
e.g. cornflakes may be harder to detect than refrigeratos. To
determine the influence of the sensor model, we performed several
experiments where we only changed the difficulty of the target
object(s). All non-target objects remain unchanged so the planner may
exploit object co-occurances with reliable objects to compensate for
the increased sensor noise.

We evaluate with DT sessions taking $b_0$ admitting between 20 and 100
abstract states with non-zero probability. We run 50 simulations in
each configuration, and have a timeout on each simulation of 30
minutes (1800 seconds)\footnote{All experiments were conducted on a
 2.66GHz Intel Xeon X5355 using one CPU core.}. The continual planning
times are reported in Figure~\ref{fig:results-time}, and the quality
data in Figure~\ref{fig:results-quality}.
%%
For each task, the goal is to find one or several objects and report
their position to a user. Usually there is a non-zero probability that
no plan exists, as the desired object might not be present in the
environment.
% MOG: I didn't do  evaluations for those, so let's remove them
% Item~$f$ reports for tasks requiring indirect sensing,
% where the robot must relocate to a room with a particular category
% (e.g. kitchen). the {\em baseline} is not able to complete Item~$f$,
% so is omitted from that reporting.  
Because reward is only allocated
once, on the achievement of the goal,  we find it
intuitive to report average plan costs and the success rates in
problems that admit a solution (i.e., positive reward scaled by a
constant factor).

@@

\begin{figure}[h!]
  % \centering
  % \includegraphics{dora1-quality}\hfill
  % \vspace{2mm}
  \includegraphics{pomdp-quality}\hfill
  % \includegraphics{pomdp-solvable-quality}\hfill
  \includegraphics{dora2-quality}\hfill
  % \vspace{2mm}
  \includegraphics{dora3-quality}\hfill
  % \vspace{2mm}
  \includegraphics{dora4-quality}\hfill
  \vspace{2mm}
  \includegraphics{dora56-quality}\hfill
  % \vspace{2mm}
  % \includegraphics{dora-cat-quality}\hfill
  \caption{Average plan costs and number of successful runs.}
  \label{fig:results-quality}
\end{figure}

We find that if sensing is reliable there is little to be gained
through DT sessions, as the greedy approach of the {\em baseline} is
sufficient. As sensing degrades contingent planning pays off.  Time
spent in DT planning increases steeply as the abstraction $\bstate_0$
becomes more refined.  That refinement seems to be paying off in terms
of the success rate, particularly for tasks $d$ and $e$. For less
refined initial configurations, the increased cost of DT
planning is partially compensated for by a decrease in Fast Downward planning
times. The relatively high success rate irrespective of the level of
refinement in the initial configuration indicates the effectiveness of
using conditional entropy to guide abstraction refinement in our
setting.

### 2011-02-08 subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
changes requested by moritz

We now describe our {\em switching} planning system that operates
according to the continual planning paradigm, given a description of
the current problem and domain in DTPDDL. The system {\em
switches}, in the sense that planning proceeds in interleaved
sessions, in which the base planner is either {\em sequential}/linear
or {\em decision-theoretic}.
%%
During a sequential session, a rewarding {\em trace} of a possible
execution is computed using a cost-optimising satisficing planner
which trades action costs, goal rewards, and determinacy.
%%
Formatted as a linear plan, the trace specifies a sequence of actions
that achieves some reachable goals following a deterministic
approximation of the problem at hand.
%%
Structurally, a trace is a sequence of elements that are either: (i) ground
actions from the DTPDDL description of the world, or (ii) atomic {\em
assumptions}, modelled as deterministic actions, made about the truth
value of facts that can only be determined at runtime (e.g., that a
box of cornflakes is located on the corner bench in the kitchen).
%%
The system always begins in a sequential session, so that plan
execution proceeds by applying DTPDDL actions from the trace in
sequence until the outcome of executing the next scheduled action is
too uncertain according to a threshold parameter~(here, 95\%). In that
eventuality a DT session commences, which tailors sensory processing
to determine whether the assumptions made in the trace hold, or which
otherwise acts to achieve the goals.

Because online DT planning in large problems is impractical (we seek
response times in seconds), a DT session plans in an abstract process
determined by the current trace and underlying belief-state. This
abstract problem is created by selecting only those propositions that
are \emph{relevant} to the subproblem the DT session is asked to
solve.
% MOG: Describe this in detail in the appropriate subsection
%  This
% abstraction is constructed by first excluding all propositions (and
% related actions) that are not true, or {\em assumed } true, of states
% in the trace, then adding them back, using as an heuristic the entropy
% of the trace assumptions conditional on a candidate
% proposition. Propositions are added, one at a time, until the number
% of states in the initial belief-state reaches a given
% threshold.\footnote{Our experiments are performed using a number of
% different threshold parameters.} 
To the resulting abstract model we
also add {\em disconfirm} and {\em confirm} actions that the
DT session can schedule in order to judge an atomic assumption
in the trace. In the abstract model these actions yield a relatively
small reward if the corresponding judgement is true (or small penalty
otherwise). If a judgement action is scheduled for execution the
DT session is terminated, and control is returned to the
sequential planner.

Finally, whether proceeding in a sequential or DT session, our
continual planner maintains a factored representation of successive
belief-states by performing belief revision. For that, we transform
the $(\pp{:init})$ declaration into a tree-shaped bayesian network
which gets updated whenever an actions is performed or an observation
is received. That belief-state representation is used:
(1) as the source of candidate determinisations for sequential
planning, (2) in determining when to switch to a DT session, and (3)
as a mechanism to guide construction of an abstract process for DT
sessions.

@@

\noindent Each assumption corresponds to one distinct {\em
relaxed} visitation of the root. Here, a conjunctive term is visited
iff its atomic subterms are visited, and zero or one of its immediate
probabilistic subterms are visited. For a sequential session, the
starting state corresponds to an assumption (usually unique) which
probability $1$. In our example, that starting state is:

@@

Encapsulating state-assumptions, we annotate the problem posed during
a sequential session with an \emph{assumptive action} $\assumptiveS{i}$ for
each element $\prob_i (T_i)$, of each probabilistic term from
$(\pp{:init})$. Here, $\assumptiveS{i}$ can be executed if no
$\assumptiveS{j}$, $j \neq i$, has been executed from the same
probabilistic term, and, either
$(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root conjunct, or
it occurs in $T_k$ for some executed $\assumptiveS{k}$.
%%
We add additional constraints that forbid the scheduling of
assumptions about facts after actions with preconditions, or effects,
that mention those facts. For example, the robot cannot assume it is
plugged into a power source immediately after it unplugs itself.
%%
Executing $\assumptiveS{i}$ in a state $\state$ effects a transition
to a successor state $\state^{T_i}$ with probability $\prob_i$, and
$\state^\bot$ with probability $1 - \prob_i$. Here, $\state^{T_i}$ is
the union of $\state$ with atomic terms from $T_i$. State
$\state^\bot$ is an added sink.

@@

%% define the value of a trace
%% $\state_0, \action_0, \state_1, \action_1,.., \state_N$ from the
%% deterministic model to equal:

%% \small
%% \begin{equation}\label{eq:tracevalue}
%% V(\state_0, \action_0, \state_1, \action_1,.., \state_N) =  \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1} \reward(\state_i, \action_i)
%% \end{equation}
%% \normalsize

%% \noindent The
%% optimal trace given a plan, i.e., the sequence of non-assumptive
%% actions from the trace, therefore has value:

@@

When an action that requires sensing is encountered, the planner
switches to a decision theoretic session. That session plans in a {\em
  small} abstract processes defined by the action that triggered the
DT session, the assumptive actions in the proceeding trace and the
current belief state. Targeted sensing is encouraged by augmenting the
reward model to reflect a heuristic value of knowing the truth about
assumptions in the trace. In detail, all rewards from the original
POMDP are retained. Additionally, for each relevant assumptive action
$\assumptiveS{i}$ in the current trace, we have a {\em disconfirm
  action} $\assumptiveDT{i}$ so that for all states $\state$:

@@

\noindent where $\$(T_i)$ (resp. $\hat\$(T_i)$) is a 
relatively\footnote{Compared to rewards from the original process
model.} small positive (negative) numeric quantity which captures the
utility the agent receives for correctly (incorrectly) rejecting an
assumption. A disconfirm action can only be executed once, and except in the
applicability of that dual corresponds to a self-transformation
%%
If \switchAction\ is the action that switched the system to a DT
session, then an assumption $\assumptiveS{i}$ is {\em relevant} if it
is directly or indirectly required for \switchAction\ to be
applicable. Continuing our simplified example, if the trace is:

@@

\noindent Taking the switching action \switchAction\ to be
$(\pp{look}~\pp{box}~\pp{office})$, we have that
$\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ is not
relevant for this session, and therefore exclude it from the
POMDP posed to the DT session.

>>>>

We now describe our {\em switching} planning system that operates
according to the continual planning paradigm, given a description of
the current problem and domain in DTPDDL. The system {\em
switches}, in the sense that planning proceeds in interleaved
sessions, in which the base planner is either {\em sequential}/linear
or {\em decision-theoretic}.
%%
During a sequential session, a rewarding {\em trace} of a possible
execution is computed using a cost-optimising satisficing planner
which trades action costs, goal rewards, and determinacy.
%%
Formatted as a linear plan, the trace specifies a sequence of actions
that achieves some reachable goals following a deterministic
approximation of the problem at hand.
%%
Structurally, a trace is a sequence of elements that are either: (i) ground
actions from the DTPDDL description of the world, or (ii) atomic {\em
assumptions}, modelled as deterministic actions, made about the truth
value of facts that can only be determined at runtime (e.g., that a
box of cornflakes is located on the corner bench in the kitchen).
%%
The system always begins in a sequential session, so that plan
execution proceeds by applying DTPDDL actions from the trace, in
sequence, until the outcome of executing the next scheduled action is
too uncertain according to a threshold parameter~(here, 95\%). In that
eventuality a DT session commences, which tailors sensory processing
to determine whether the assumptions made in the trace hold, or which
otherwise acts to achieve the goals.

Because online DT planning in large problems is impractical (we seek
response times in seconds), a DT session plans for an abstract process
determined by the current trace and underlying belief-state. That
abstract problem is characterised by a limited number of propositions,
chosen due to their {\em relevance} to the action whose scheduled
execution triggered the switch to the DT session.
%%the subproblem the DT session is asked tosolve.
% MOG: Describe this in detail in the appropriate subsection
%  This
% abstraction is constructed by first excluding all propositions (and
% related actions) that are not true, or {\em assumed } true, of states
% in the trace, then adding them back, using as an heuristic the entropy
% of the trace assumptions conditional on a candidate
% proposition. Propositions are added, one at a time, until the number
% of states in the initial belief-state reaches a given
% threshold.\footnote{Our experiments are performed using a number of
% different threshold parameters.} 
To that abstract model we also add {\em disconfirm} and {\em confirm}
actions that the DT session can schedule in order to judge an atomic
assumption in the trace. Those actions yield a relatively small reward
if the corresponding judgement is true (or small penalty
otherwise). If a judgement action is scheduled for execution the DT
session is terminated, and control is returned to a new sequential
session.

Whatever the session type, our continual planner maintains a factored
representation of successive belief-states by performing belief
revision. As an internal representation of the $(\pp{:init})$
declaration, we keep a tree-shaped Bayesian network which gets updated
whenever an action is performed, or an observation received. That
belief-state representation is used: (1) as the source of candidate
determinisations for sequential planning, (2) in determining when to
switch to a DT session, and (3) as a mechanism to guide construction
of an abstract process for DT sessions.

@@

\noindent Each assumption corresponds to one distinct {\em
relaxed} visitation of the root term. Here, a conjunctive term is visited
iff its atomic subterms are visited, and zero or one of its immediate
probabilistic subterms are visited. For a sequential session, the
starting state corresponds to an assumption (usually unique) with
probability $1$. In our example, that starting state is:

@@

Encapsulating state-assumptions, we annotate the problem posed during
a sequential session with an \emph{assumptive action} $\assumptiveS{i}$ for
each element $\prob_i (T_i)$, of each probabilistic term from
$(\pp{:init})$. Here, $\assumptiveS{i}$ can be executed if no
$\assumptiveS{j}$, $j \neq i$, has been executed from the same
probabilistic term, and, either
$(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root conjunct, or
it occurs in $T_k$ for some executed $\assumptiveS{k}$.
%%
We also add constraints that forbid scheduling of
assumptions about facts after actions with preconditions or effects
that mention those facts. For example, the robot cannot assume it is
plugged into a power source immediately after it unplugs itself.
%%
Executing $\assumptiveS{i}$ in a state $\state$ effects a transition
to a successor state $\state^{T_i}$ with probability $\prob_i$, and
$\state^\bot$ with probability $1 - \prob_i$. Here, $\state^{T_i}$ is
the union of $\state$ with atomic terms from $T_i$. State
$\state^\bot$ is an added sink.

@@

@@

When an action is scheduled whose outcome is uncertain according to
the underlying belief-state, the planner switches to a DT
session. That plans for {\em small} abstract processes defined
according to the action that triggered the DT session, the assumptive
actions in the proceeding trace, and the current
belief-state. Targeted sensing is encouraged by augmenting the reward
model to reflect a heuristic value of knowing the truth about
assumptions. In detail, all rewards from the underlying problem are
retained. Additionally, for each relevant assumptive action
$\assumptiveS{i}$ in the current trace, we have a {\em disconfirm
action} $\assumptiveDT{i}$ so that for all states $\state$:

@@

\noindent where $\$(T_i)$ (resp. $\hat\$(T_i)$) is a 
relatively small positive (negative) numeric quantity\footnote{Small
relative to rewards from the original process model.} which captures
the utility the agent receives for correctly (incorrectly) rejecting
an assumption. A disconfirm action can only be executed once, and
except in the applicability of that dual corresponds to a
self-transformation
%%
We only consider {\em relevant} assumptions and their duals when
constructing the abstract model.  If \switchAction\ is the action that
switched the system to a DT session, then an assumption
$\assumptiveS{i}$ is {\em relevant} if it is necessary for the outcome
of \switchAction\ to be determined. Continuing our simplified example,
if the trace is:

@@

\noindent Taking the switching action \switchAction\ to be
$(\pp{look}~\pp{box}~\pp{office})$, we have that
$\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ is not
relevant for this session, and therefore exclude it from the
abstract decision process.
